### 2016-08-25 section/2_workless.tex
shrink through section 3 and 4

\section{\Tool design}
\label{sec:design}
\Tool consists of a series of analysis that
judges whether a given loop belongs to any root cause type discussed in
Section \ref{sec:study_tax}.
Its design
follows the following principles.
\begin{itemize}
\item Failure diagnosis, not bug detection. \Tool will be  
used together with other performance diagnosis tools \cite{SongOOPSLA2014}
and focus on a small
number of loops that are most correlated with a specific performance symptom,
instead of being applied to the whole program. Therefore, we will have different
design trade-offs in terms of coverage and accuracy, comparing with 
bug detection tools.

\item Static-dynamic hybrid analysis. Static analysis alone
will not be able to provide all the needed information to judge whether
a loop is inefficient. However, 
dynamic analysis alone will incur too much overhead.
Therefore, we use a hybrid approach throughout our design.

\item Sampling. Loop-related 
performance problems have the unique nature of repetitiveness, which make 
them a natural fit for random sampling. We will design different
sampling schemes for different analysis.
\end{itemize}

@@

Our resultless checker includes two parts. First, we use static analysis
to figure out which are the side-effect instructions in a loop and hence
decide whether a loop belongs to 0*, 0*1?, [0$|$1]*, or 1*. Second, to
accurately determine 0*1? and [0$|$1]* loops, we use
dynamic analysis to figure out what portion of loop iterations are
resultless at run time.
%, which will help decide whether the loop is indeed
%inefficient.

@@

\subsubsection{Static Analysis}
\label{sec:s_workless}

\Tool considers two types of instructions as side-effect instructions:
(1) write to heap or global variables;
(2) write to a stack variable, defined outside the loop, and the update value
may be used after the loop. The second condition is checked through liveness
analysis.
\Tool also analyzes all functions called
by a loop directly or indirectly --- a function $F$ that updates variables
defined outside $F$ makes the corresponding call statement in $F$'s
caller a side-effect instruction.
We consider all calls to library functions or through function pointers 
as having side effects, 
unless the library functions are specially marked by us in a white list.
This analysis is complete, but not sound. We may identify side-effect
instructions that are actually not.

\Tool then categorizes loops into four types.
Given a natural loop $l$, when $l$
contains at least one side-effect instruction along every path that
starts from the loop header and ends at the loop header, it is a 1* loop;
if there exists at least one side-effect instruction inside $l$, but not on
every path,
it is a [0$|$1]* loop; if there is no side-effect instructions inside $l$,
$l$ is either a 0* or a 0*1? loop.
In the last case, \Tool further checks all the loop exit-blocks. If there 
exists an exit-block of $l$ that contains a side-effect instruction and is  
post-dominated by another exit-block of $l$, this is a 0*1? loop.
For example, line 5 in Figure \ref{fig:Mozilla347306} is such an exit block.
Otherwise, $l$ is a 0* loop.

Note that since the 1* pattern contains the least amount of information
about computation \textit{inefficiency}, \Tool will not report a loop's
root-cause type as 1*, if more informative root-cause type is identified 
for this loop (e.g., cross-iteration or cross-loop redundancy).

%TODO Shan will rewrite the next two paragraphs
\subsubsection{Dynamic Monitoring}
\label{sec:d_workless}

Except for 0*, none of the other three type of loops are inefficient for sure.
We need dynamic analysis to figure out what portion of loop iterations are
resultless at run time, which will help decide whether the loop worths fixing.

For a 0*1? loop, since it only generates results in the last iteration, we 
only need to know the total number of loop iterations 
to figure out the loop
\textit{resultful rate}, which we define as the ratio between
the number of iterations with
side effects and the total number of iterations. 
The implementation is straightforward
--- we initialize a local counter to be 0 in the pre-header of the loop; we 
increase the counter by 1 in the loop header to count the number of 
iterations; we dump that counter to log when the loop exits.

For [0$|$1]*, we need to count not only the total number of iterations, but
also the exact number of iterations that execute side-effect instructions
at run time. 
To do that, our instrumentation uses a local boolean variable 
\texttt{HasResult} to represent whether one iteration have side effect or not. 
\texttt{HasResult} is set to \texttt{False} in the loop header, and set to
\texttt{True} after each side-effect instruction. It will be used to help
count the number of side-effect iterations. For performance concerns,
before instrumenting side-effect blocks, we check whether there are 
post-domination relation between each pair of side-effect blocks. 
If both block A and block B are side-effect blocks and block A post-dominates 
block B, we only instrument block A to update \texttt{HasResult}.

>>>>

\section{\Tool design}
\label{sec:design}
The design of \Tool follows three principles.
First, \textit{symptom oriented}. \Tool will be  
used together with other performance diagnosis tools \cite{SongOOPSLA2014}
or profilers,
and focus on a small
number of loops that are most correlated with a specific already-observed 
performance symptom,
instead of the whole program. Therefore, we will have different
design trade-offs in terms of coverage and accuracy, comparing with 
bug detection tools.
Second, \textit{static--dynamic hybrid}. Static analysis alone
cannot provide all the needed information to judge whether
a loop is inefficient. However, 
dynamic analysis alone will incur too much overhead.
Therefore, we use a hybrid approach throughout our design.
Third, \textit{sampling}. Loop-related 
performance problems have the unique nature of repetitiveness, which make 
them a natural fit for random sampling. We will design different
sampling schemes for different analysis.

We envision \Tool to be used in the following way.
Existing profilers and statistical debugging tool first
identify suspicious loops that are most correlated with 
certain performance 
symptoms \cite{SongOOPSLA2014}. 
\Tool will then apply a series of analysis to judge whether a given
loop belongs to any root-cause type discussed in Section \ref{sec:study}
and recommend corresponding fix strategy. This series of analysis could be
conducted during either in-house diagnosis, where failure-triggering inputs
are available, or production runs, with
the run-time analysis component of \Tool running at user sites.

@@

Our resultless checker includes two parts. First, we use static analysis
to figure out which are the side-effect instructions in a loop and hence
decide whether a loop might belong to 0*, 0*1?, [0$|$1]*, or 1*. Second, to
determine whether 0*1? and [0$|$1]* loops are indeed inefficient
under specific workload, we use
dynamic analysis to figure out what portion of loop iterations are
resultless at run time.
%, which will help decide whether the loop is indeed
%inefficient.

@@

\subsubsection{Static analysis}
\label{sec:s_workless}

\Tool considers two types of instructions as side-effect instructions:
(1) write to heap or global variables;
(2) write to a stack variable, defined outside the loop, and the update value
may be used after the loop. The second condition is checked through liveness
analysis.
\Tool also analyzes all functions called
by a loop directly or indirectly --- a function $F$ that updates variables
defined outside $F$ makes the corresponding call statement in $F$'s
caller a side-effect instruction.
\comment{
We consider all calls to library functions or through function pointers 
as having side effects, 
unless the library functions are specially marked by us in a white list.
}
This analysis is complete, but not sound. We may identify side-effect
instructions that are actually not. %to be moved to threats to validity

\Tool then categorizes loops into four types.
Given a natural loop $l$, when $l$
contains at least one side-effect instruction along every path that
starts from the loop header and ends at the loop header, it is a 1* loop;
if there exists at least one side-effect instruction inside $l$, but not on
every path,
it is a [0$|$1]* loop; if there is no side-effect instructions inside $l$,
$l$ is either a 0* or a 0*1? loop, which will be differentiated by checking
all the loop exit-blocks.
%If there 
%exists an exit-block of $l$ that contains a side-effect instruction and is  
%post-dominated by another exit-block of $l$, this is a 0*1? loop.
For example, the loop in Figure \ref{fig:Mozilla347306} is 0*1?, because 
its exit block on Line 5 contains side effects.

\comment{
Note that since the 1* pattern contains the least amount of information
about computation \textit{inefficiency}, \Tool will not report a loop's
root-cause type as 1*, if more informative root-cause type is identified 
for this loop (e.g., cross-iteration or cross-loop redundancy).
}

\subsubsection{Dynamic monitoring}
\label{sec:d_workless}

Further checking is conducted to determine whether 0*1? and [0$|$1]* 
loops are inefficient.

For a 0*1? loop, since it only generates results in the last iteration, we 
only need to know the total number of loop iterations of each loop instance
to figure out the loop
\textit{resultful rate}, which we define as the ratio between
the number of iterations with
side effects and the total number of iterations. 
The implementation is straightforward
--- we initialize a local counter to be 0 in the pre-header of the loop; we 
increase the counter by 1 in the loop header to count the number of 
iterations; we dump that counter to log when the loop exits.

For [0$|$1]*, we need to count not only the total number of iterations, but
also the exact number of iterations that execute side-effect instructions
at run time. 
To do that, our instrumentation uses a local boolean variable 
\texttt{HasResult} to represent whether one iteration have side effect or not. 
\texttt{HasResult} is set to \texttt{False} in the loop header, and set to
\texttt{True} after each side-effect instruction. It will be used to help
count the number of side-effect iterations. 
\comment{
For performance concerns,
before instrumenting side-effect blocks, we check whether there are 
post-domination relation between each pair of side-effect blocks. 
If both block A and block B are side-effect blocks and block A post-dominates 
block B, we only instrument block A to update \texttt{HasResult}. 
}
